---
title: "Week 2 Control Structures"
author: "Me"
date: "6/6/2020"
output: html_document
---


If/else intro
```{r}
x <- runif(1, 0, 10)
if(x > 3) {
  y <- 10
} else {
  y <- 0
}
print(c(x,y))
```
Alternate way to write above if/else command

```{r}
x <- runif(1, 0, 10)
y <- if(x > 3) {
  10
} else {
  0
}
print(x )
```

For loops print list:

```{r}
for (i in 1:10) {
  print(i)
}
```
For loops print the elements at list indexes:

```{r}
x <- c('a' ,'b', 'c', 'd')

for (i in 1:4) {
  print(x[i])
}
```

How to print each element of a list:

```{r}
for(letter in x) {
  print(letter)
}
```

Curly braces aren't necessary for one-line for loops:

```{r}
for(i in 1:4) print(x[i])
```

Nested for loops:

```{r}
x <- matrix(1:6, 2, 3)
print(x)

for (i in seq_len(nrow(x))) {
  for(j in seq_len(ncol(x))) {
    print(x[i,j])
  }
}
```

While Loops:

```{r}
count <- 0
while (count < 10) {
  print(count)
  count <- count + 1
}
```

repeat Loops - repeat loops run forever until a break statement is encountered

```{r}
x0 <- 1
tol <- 1e-8

repeat {
  x1 <- 2
  
  if (abs(x1 - x0) < tol) {
    break
  } else {
    x0 <- x1
  }
}
```

next, break:
*Next is used to skip one iteration in a loop

```{r}

for(i in 1:30) {
  if(i %% 2 ==0) {
  next
  }
  print(i)
}
```

























