---
title: "Loop Functions for Command Line"
author: "Me"
date: "6/11/2020"
output: html_document
---

****************************************************************
lapply #1 - note that when using one function as an argument to lapply, the parantheses are not required

```{r}
x <- list(a = 1:5, b = rnorm(10))
x
lapply(x, mean)
```
****************************************************************
lapply #2 - can use it to generate variable lengths of numbers

```{r}
x = 1:4
lapply(x, runif)

lapply(x, runif, min = 0, max = 10)
```
****************************************************************
lapply #3 - applying an anonymous function:

```{r}
x <- list(a = matrix(1:4, 2, 2), b = matrix(1:6, 3, 2))
lapply(x, function(elt) {elt[,1]})



```
****************************************************************
sapply #1 - Like lapply but sapply tries to simplify the result of lapply(). 
F

```{r}
x <- list(a = 1:4, b = rnorm(10), c = rnorm(20, 1), d = rnorm(100, 5))
sapply(x, mean)
```
****************************************************************
split #1 - Takes a vector or other object and splits it into groups 

```{r}
str(split)
```
^^^ 
-x is a vector, list or dataframe
-f is a factor or a list of factors
-drop indicates if empty factors should be dropped

*bonus - the gl function is used for mapping level and pairs well with the split function.

```{r}
x <- c(rnorm(10), runif(10), rnorm(10, 1))
f <- gl(3, 10)
split(x,f)

```

Split #2
Can add further complexity to the code by taking the mean on the same line:

```{r}
lapply(split(x, f), mean)

```
*** Cool use of split on a large data set***
Splits airquality dataframe by month


```{r}
library(datasets)
head(airquality)
s <- split(airquality, airquality$Month)
str(s)
```

Split #3:
Combing sapply and split for a cleaner output

```{r}
sapply(s, function(x) {
    colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE)
})
```
***********************************************************
Interactions of the gl function

```{r}
x <- rnorm(10)
f1 <- gl(2,5)
f2 <- gl(5,2)
f1
f2
```
Now use interact function to combine f1 and f2

```{r}
interaction(f1, f2)
```

**********************************************
tapply - application of a function onto a subset of vectors


```{r}
str(tapply)
```

```{r}
set.seed(1)
x <- c(rnorm(100), runif(100), rnorm(100, 1))
f <- gl(3, 100)
tapply(x, f, mean)
```
***********************************************
apply() - function to evaluate a function (usually anonymous) over the margins of an array. For columns, set margin to 2. For rows, set margin to 1.

```{r}
str(apply)
```

```{r}
x <- matrix(rnorm(200, 20, 20),20,10)
apply(x,2,mean)
```

```{r}
apply(x, 1, mean)
```

Quantiles2 with apply()

```{r}
x <- matrix(rnorm(200), 20, 10)
apply(x,1, quantile, probs = c(0.25, 0.75))
```

**********************************************
mapply() multivariat apply to iterate over multipl R objects

```{r}
str(mapply)
```
***The two functions below are equivalent
```{r}
list(rep(1,4), rep(2,3), rep(3,2), rep(4,1))

mapply(rep, 1:4, 4:1)


```

mapply#2

```{r}
noise <- function(n, mean, sd){
    rnorm(n, mean, sd)
}
noise(5,1,2)
noise(1:5, 1:5, 2)
```
Problem: noise does not read lists well for its argument -> enter mapply

```{r}
mapply(noise, 1:5, 1:5, 2)
```

Mapply finds the proper lists for each iteration. This short script would be the same as writing out: list(noise(1, 1, 2), noise(2, 2, 2), noise(3, 3, 2), noise(4, 4, 2), noise(5, 5, 2))






















